---
title: "Repeat Sampling"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown empirically highlights the problem with drawing conclusions (or inference) through repeated sampling of the same dataset.


Both models are drawn randomly from the same distribution
```{r}
model1 = rnorm(500)
model2 = rnorm(500)
```

There is a slight difference in means.
```{r}
(c(mean(model1), mean(model2)))
```

Let's sample from the models and calculate the mean for each sample
```{r}
n_samples = 1000
sample_size = 200
sample_mean_fn = function(arr) mean(sample(arr, sample_size))

mns1 = sapply(1:n_samples, function(x) sample_mean_fn(model1))
mns2 = sapply(1:n_samples, function(x) sample_mean_fn(model2))
```


We see the plot rapidly becomes statistically significant (a y-axis value of 1.3 corresponds to a p-value of .05), even though both models are drawn from the same population. This is because the t-test assumes independence between samples. However, the samples are continually drawn from the same 2 datasets. Therefore, they are not independent. The t-test is "fooled" into thinking the small effects are significant.
```{r}
plot(2:n_samples, 
     sapply(2:n_samples, function(i) 
       -log10(t.test(mns1[1:i], mns2[1:i])$p.value)), 
     ylab = '-log p-value', 
     xlab = '# experiments')
```

What about the non-parametric Wilcoxon rank test?

We will see the same result. It's not an issue of parameters.
```{r}
plot(2:n_samples, 
     sapply(2:n_samples, function(i) 
       -log10(wilcox.test(mns1[1:i], mns2[1:i])$p.value)), 
     ylab = '-log p-value', 
     xlab = '# experiments')
```

However, the distribution of means shows significant overlap.
```{r}
library(tidyverse)
model1_df = data.frame(mean = mns1, model = "model1")
model2_df = data.frame(mean = mns2, model = "model2")
rbind(model1_df, model2_df) %>% 
  ggplot(aes(x=mean)) + 
  geom_histogram(aes(color = model, fill=model, y = ..density..)) + 
  geom_density() +
  facet_grid(. ~ model)
```

